---
title: "Data Processing"
author: "Giorgos Tsalidis"
date: "1/31/2020"
output: html_document
---

### Feature Selection

In this section we will use the Correlation Feature Selection (CFS) method to select attributes
to be used when training and testing the machine learning models.
```{r}
data = iris
# split into training and validation datasets
set.seed(1234)
ind <- sample(2, nrow(data), replace=TRUE, prob=c(0.7, 0.3))
trainData <- data[ind==1,]
validationData <- data[ind==2,]
# keep only instances that do not have missing values.
trainData <- trainData[complete.cases(trainData),]
validationData <- validationData[complete.cases(validationData),]
```

### Using the CFS method

```{r}
library(FSelector)
subset <- cfs(Species ~ ., trainData)
f <- as.simple.formula(subset, "Species")
print(f)
```
The output is a formula that we can use in various classification algorithms and says that
according to the CFS algorithm and the training dataset, the best features in order to predict
the Species of the Iris flowers are the Petal Length and the Petal Width.

For example in the Naive Bayes algorithm, we will use both the formula that includes all the
attributes for predicting the Species and the formula derived from CFS.
```{r}
library(e1071)
model <- naiveBayes(Species ~ ., data=trainData, laplace = 1)
simpler_model <- naiveBayes(f, data=trainData, laplace = 1)
pred <- predict(model, validationData)
simpler_pred <- predict(simpler_model, validationData)
library(MLmetrics)
train_pred <- predict(model, trainData)
train_simpler_pred <- predict(simpler_model, trainData)
paste("Accuracy in training all attributes",
Accuracy(train_pred, trainData$Species), sep=" - ")
paste("Accuracy in training CFS attributes",
Accuracy(train_simpler_pred, trainData$Species), sep=" - ")
paste("Accuracy in validation all attributes",
Accuracy(pred, validationData$Species), sep=" - ")
paste("Accuracy in validation CFS attributes",
Accuracy(simpler_pred, validationData$Species), sep=" - ")
```

## Wrapper Methods
### Forward Search

```{r}
# Downloading the file
fileURL <- "http://archive.ics.uci.edu/ml/machine-learning-databases/breast-cancer-wisconsin/breast-cancer-wisconsin.data"
#download.file(fileURL, destfile="../data/breast-cancer-wisconsin.data")
data <- read.table('../data/breast-cancer-wisconsin.data', na.strings = "?", sep=",")
data <- data[,-1]
names(data) <- c("ClumpThickness",
"UniformityCellSize",
"UniformityCellShape",
"MarginalAdhesion",
"SingleEpithelialCellSize",
"BareNuclei",
"BlandChromatin",
"NormalNucleoli",
"Mitoses",
"Class")
data$Class <- factor(data$Class, levels=c(2,4), labels=c("benign", "malignant"))
set.seed(1234)
ind <- sample(2, nrow(data), replace=TRUE, prob=c(0.7, 0.3))
trainData <- data[ind==1,]
validationData <- data[ind==2,]
# remove cases with missing data
trainData <- trainData[complete.cases(trainData),]
validationData <- validationData[complete.cases(validationData),]
```
Apply Kfold cross validation

```{r}
library(FSelector)
library(rpart)
evaluator <- function(subset) {
#k-fold cross validation
k <- 5
splits <- runif(nrow(trainData))
results = sapply(1:k, function(i) {
  test.idx <- (splits >= (i - 1) / k) & (splits < i / k)
  train.idx <- !test.idx
  test <- trainData[test.idx, , drop=FALSE]
  train <- trainData[train.idx, , drop=FALSE]
  tree <- rpart(as.simple.formula(subset, "Class"), train)
  error.rate = sum(test$Class != predict(tree, test, type="c")) / nrow(test)
return(1 - error.rate)
})
print(subset)
print(mean(results))
return(mean(results))
}
subset <- forward.search(names(trainData)[-10], evaluator)
```


```{r}
f <- as.simple.formula(subset, "Class")
print(f)
```

The fact that we performed forward search using decision trees in order to get a formula with
a subset of attributes, doesn’t stop us from using another classification model for training and
prediction. For example as in the previous examples in tehi chapter, we can use the Naive Bayes
algorithm to evaluate the forward selection algorithm both in the training and the validation
datasets under the accuracy metric.

```{r}
library(e1071)
model <- naiveBayes(Class ~ ., data=trainData, laplace = 1)
simpler_model <- naiveBayes(f, data=trainData, laplace = 1)
pred <- predict(model, validationData)
simpler_pred <- predict(simpler_model, validationData)
library(MLmetrics)
train_pred <- predict(model, trainData)
train_simpler_pred <- predict(simpler_model, trainData)
paste("Accuracy in training all attributes",
Accuracy(train_pred, trainData$Class), sep=" - ")
paste("Accuracy in training forward search attributes",
Accuracy(train_simpler_pred, trainData$Class), sep=" - ")
paste("Accuracy in validation all attributes",
Accuracy(pred, validationData$Class), sep=" - ")
paste("Accuracy in validation forward search attributes",
Accuracy(simpler_pred, validationData$Class), sep=" - ")
```

## Dimensionality Reduction
### Principal Components Analysis

Step 1: Data
First we are going to make a dataset and plot it.
```{r}
d <- c(2.5, 2.4, 0.5, 0.7, 2.2, 2.9, 1.9, 2.2, 3.1, 3.0, 2.3,
2.7, 2, 1.6, 1, 1.1, 1.5, 1.6, 1.1, 0.9)
data <- matrix(d, ncol=2, byrow = T)
plot(data, xlab="x1", ylab="x2")
```

Step 2: Subtract the mean

We subtract the mean of each attribute (x1, x2) from the respective values (centering the data)
using the function scale .

```{r}
data_norm <- scale(data, scale=F)
plot(data_norm, xlab="x1", ylab="x2")
```
Step 3: Calculate the covariance matrix

We calculate the covariance matrix using the function cov .

```{r}
S <- cov(data_norm)
print(S)
```
Step 4: Computer the eigenvectors of the covariance matrix

Using the svd method (Singular Value Decomposition) we derive the eigenvectors of the
covariance matrix. First we print the decomposition:

```{r}
udv <- svd(S)
print(udv)
```

then we plot the eigenvectors:
```{r}
plot(data_norm, asp=1, xlab="x1", ylab="x2")
arrows(0,0,udv$u[1,1],udv$u[2,1], lwd=1)
arrows(0,0,udv$u[1,2],udv$u[2,2], lwd=0.5)
```

Step 5: Choosing components

We can use a barplot to display the variance accounted for each component.
```{r}
barplot(udv$d)
print(cumsum(udv$d)/sum(udv$d))
```
We can see that the 1st component accounts for more than 95% of the variance.

Step 6: Picking the 1st component

We transform the 2D dataset into a 1D dataset using just the 1st Principal Component (PC).

```{r}
data_new <- t(udv$u[,1,drop=FALSE]) %*% t(data_norm)
data_new

plot(data_new,data_new,asp=1,xlab="x1", ylab="x2")
```
### PCA using an R function

In this subsection, we will use the prcomp function that performs PCA in one step. We will apply
it on the Iris dataset and test the gains or losses in terms of accuracy using the kNN algorithm.
```{r}
# download the file
data = iris
set.seed(1234)
ind <- sample(2, nrow(data), replace=TRUE, prob=c(0.7, 0.3))
trainData <- data[ind==1,]
validationData <- data[ind==2,]
library(class)
library(stats)
trainData <- trainData[complete.cases(trainData),]
validationData <- validationData[complete.cases(validationData),]
trainDataX <- trainData[,-ncol(trainData)]
logTrainDataX <- log(trainDataX)
train.pca <- prcomp(logTrainDataX, center = TRUE, scale. = TRUE)
summary(train.pca)
```


```{r}
trainDataY <- trainData$Species
validationDataX <- validationData[,-ncol(trainData)]
# Let's also transform the validation data
logValidationDataX <- log(validationDataX)
validation.pca <- predict(train.pca, newdata=logValidationDataX)
validationDataY <- validationData$Species
# no pca prediction
prediction = knn(trainDataX, validationDataX, trainDataY, k = 3)
# So let's predict using only the 7 principal components
prediction_pca = knn(train.pca$x[,1:2], validation.pca[,1:2], trainDataY, k = 3)
cat("Confusion matrix:\n")
xtab = table(prediction, validationData$Species)
print(xtab)
cat("\nEvaluation:\n\n")
accuracy = sum(prediction == validationData$Species)/length(validationData$Species)
cat(paste("Accuracy:\t", format(accuracy, digits=2), "\n",sep=" "))
cat("Confusion matrix PCA:\n")
xtab = table(prediction_pca, validationData$Species)
print(xtab)

cat("\nEvaluation PCA:\n\n")
accuracy = sum(prediction_pca == validationData$Species)/length(validationData$Species)
cat(paste("Accuracy PCA:\t", format(accuracy, digits=2), "\n",sep=" "))
```

Using the PCA we can also plot in 2 dimensions, high-dimensional data, by using just the first
two components.

```{r}
plot(train.pca$x[trainData$Species == 'setosa',1:2], col="blue", ylim = c(-3, 3),
xlim=c(-3,3), asp=1)
points(train.pca$x[trainData$Species == 'versicolor',1:2], pch = 3, col="red")
points(train.pca$x[trainData$Species == 'virginica',1:2], pch = 4, col="green")
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```

